\chapter{Related technologies}
%------------------------------------------------------------------------------
%10 pages. For motivation: Design pattern figure. 
%------------------------------------------------------------------------------

The current chapter introduces the different technologies used to build the
\ptool{}. In \autoref{sect:javabasics} we give a short overview about the the
Java platform and the internals of the Java binaries. In
\autoref{sect:serverside} we go through the frameworks which were used to build
and implement the server side of the tool. Last, but not least
\autoref{sect:clientside} summarizes the technologies used on the client side.
At some places, the followings rely on official tool documentation as marked by citations in
section titles.

\section{Java basics}\label{sect:javabasics}
 
\subsection{Java byte code specification~\cite{JavaSpec}}
% Java platform
The Java programming language is a general-purpose, concurrent,
object-oriented language. Its syntax is similar to C and C++, but it omits many
of the features that make C and C++ complex, confusing, and unsafe. The Java
platform was initially developed to address the problems of building software
for networked consumer devices. It was designed to support multiple host
architectures and to allow secure delivery of software components. To meet these
requirements, compiled code had to survive transport across networks, operate on
any client, and assure the client that it was safe to run.

% Java Virtual machine
The Java virtual machine is an abstract computing machine. Like a real computing
machine, it has an instruction set and manipulates various memory areas at
runtime. It is reasonably common to implement a programming language using a
virtual machine. The Java virtual machine knows nothing of the Java programming
language, only of a particular binary format, the  \code{class} file format. A
\code{class} file contains Java virtual machine instructions (or 
\emph{bytecodes}) and a symbol table, as well as other ancillary information.

% What is a class file
The class file format is a hardware- and operating system-independent binary
format which is executed by the virtual machine. Typically (but not necessarily)
stored in a file. The class file format precisely defines the representation of a
class or interface, including details such as byte ordering that might be taken
for granted in a platform-specific object file format.

% Compile source code
\picr{compilation.pdf}{Compilation and execution of Java classes}
\autoref{fig:compilation.pdf} illustrates the procedure of compiling and
executing a Java class: The source file is compiled into a Java class file,
loaded by the byte code interpreter and executed. In order to implement
additional features, researchers may want to transform class files (drawn with
bold lines) before they get actually executed.

\subsection{Internals of a Java class file~\cite{BCEL}}
% Internals of the class file
\pic{classfile.pdf}{Simplified structure of a class file}
\autoref{fig:classfile.pdf} shows a simplified example of the contents of a
Java class file: It starts with a header containing a "magic number"
(0xCAFEBABE) and the version number, followed by the constant pool, which can be
roughly thought of as the text segment of an executable, the access rights of
the class encoded by a bit mask, a list of interfaces implemented by the class,
lists containing the fields and methods of the class, and finally the class
attributes, e.g., the SourceFile attribute telling the name of the source file.
Attributes are a way of putting additional, user-defined information into class
file data structures. For example, a custom class loader may evaluate such
attribute data in order to perform its transformations. The JVM specification
declares that unknown, i.e., user-defined attributes must be ignored by any
Virtual Machine implementation.

Because all of the information needed to dynamically resolve the symbolic
references to classes, fields and methods at run-time is coded with string
constants, the constant pool contains in fact the largest portion of an average
class file, approximately 60\%. In fact, this makes the constant pool an easy
target for code manipulation issues. The byte code instructions themselves just
make up 12\%. 

The right upper box shows a "zoomed" excerpt of the constant pool, while the
the box on the lower right depicts some instructions that are contained within a method
of the example class. These instructions represent the straightforward
translation of the well-known \code{System.out.println("Hello, world");} statement.

The first instruction loads the contents of the field out of class
\code{java.lang.System} onto the operand stack. This is an instance of the class
\code{java.io.PrintStream}. The \code{ldc} ("Load constant") pushes a reference
to the string "Hello world" on the stack. The next instruction invokes the
instance method \code{println} which takes both values as parameters (Instance
methods always implicitly take an instance reference as their first argument).

Instructions, other data structures within the class file and constants
themselves may refer to constants in the constant pool. Such references are
implemented via fixed indexes encoded directly into the instructions. This is
illustrated for some items of the figure emphasized with a surrounding box.

For example, the \code{invokevirtual} instruction refers to a MethodRef constant that
contains information about the name of the called method, the signature (i.e.,
the encoded argument and return types), and to which class the method belongs.
In fact, as emphasized by the boxed value, the \code{MethodRef} constant itself just
refers to other entries holding the real data, e.g., it refers to a
\code{ConstantClass} entry containing a symbolic reference to the class
\code{java.io.PrintStream}. To keep the class file compact, such constants are
typically shared by different instructions and other constant pool entries.
Similarly, a field is represented by a \code{Fieldref} constant that includes
information about the name, the type and the containing class of the field.

The constant pool basically holds the following types of constants: References
to methods, fields and classes, strings, integers, floats, longs, and doubles.


\section{Server side}\label{sect:serverside}

\subsection{Application platform}

\subsubsection{Spring Framework~\cite{SpringWeb}}
%
The Spring Framework is an open source application framework and Inversion of
Control container for the Java platform. The core features of the Spring
Framework can be used by any Java application, but there are extensions for
building web applications on top of the Java EE platform. Although the Spring
Framework does not impose any specific programming model, it has become popular
in the Java community as an alternative to, replacement for, or even addition to
the Enterprise JavaBean (EJB) model.

%
The main advantage of using Spring framework is a configurability through
dependency injection. Using its features it is a solid platform for creating
clear, well-tested and maintainable applications.


\subsubsection{Oracle database~\cite{Oracle11g}}
The Oracle Database (also known as Oracle DB, Oracle RDBMS or just Oracle) is a
relational database management system (RDBMS) from the Oracle Corporation.
Originally developed in 1977 by Lawrence Ellison and other developers, Oracle DB
is one of the most trusted and widely-used relational database engines.

The system is built around a relational database framework in which data objects
may be directly accessed by users (or an application front end) through
structured query language (SQL). Oracle is a fully scalable relational database
architecture and is often used by global enterprises, which manage and process
data across wide and local area networks. The Oracle database has its own
network component to allow communications across networks.


\subsection{Build system}

\subsubsection{Maven~\cite{ApacheMaven}}
In three words Apache Maven is a project management framework. The main goal is 
similar to Ant but Maven offers an integrated approach to make Java development
manageable. 

Maven standardizes the entire development process by defining explicit lifecycle
steps for dependency management, building, packaging, deploying and so on.
Although the lifecycle elements are set, they can be configurable through the 
key element of Maven: the Project Object Model (POM). POM is an XML document in 
the project's folder describing every necessary information about it to make 
maven be able to do its job. 

Through the POM file almost everything related to the project is configurable,
yet the most important feature is the dependency management. One can define
certain dependencies in this file, which will be download automatically from 
the Maven repositories.

Also POM is the place to define almost every relevant information about the project:
source and binary folders, package format, deployment location, etc. These settings
get applied during the build process.

Maven is also extendable through plugin mechanism. One can develop extensions for it
to achieve various non-standard goals such as integration testing or building 
non-standalone Java applications.

\subsubsection{Tycho~\cite{TychoWeb}}
Tycho is a set of Maven plugins and extensions for building Eclipse plugins and
OSGi bundles with Maven. It exists because Eclipse plugins have their own way of
describing metadata such as version numbers and dependencies, which normally
would be placed in the Maven POM. Tycho reuses this native metadata 
and uses the POM to configure and drive the build. Tycho also knows
how to run JUnit test plugins using OSGi runtime and there is also support for
sharing build results using Maven artifact repositories. Tycho introduces
new packaging types and the corresponding lifecycle bindings that allow Maven to
use OSGi and Eclipse metadata during a Maven build.

Although Tycho is useful for building Eclipse plugins in a headless way but it
still has not reached the production quality (the current version number is
0.16) and it is still under heavy development. The current modifications and
enchantments can be found on the project's website.

\subsubsection{Apache Commons Byte Code Engineering Library~\cite{BCEL}} 
% what is apache bcel
The Byte Code Engineering Library (BCEL) is intended to give users a
convenient way to analyze, create, and manipulate (binary) Java files. Classes
are represented by objects which contain all the symbolic information of the
given class: methods, fields and byte code instructions, in particular.

% static part of the api
The BCEL API abstracts from the concrete circumstances of the Java Virtual
Machine and how to read and write binary Java class files. The package 
responsible for reading the class files is called the ''static'' part.
It contains a set of classes reflecting the class file format not intended for
byte code modifications. The classes may be used to read and write class files
from or to a file. This is useful especially for analyzing Java classes without
having the source files at hand. The main data structure is called
\code{JavaClass} which contains methods, fields, etc..

% what are the benefits of using this
By using the BCEL library we can extract all static information from the Java binaries without
ending up doing parsing binary data by hand. With this library it fairly easy to
(i) gather any structural information such as class name, superclass, signature
of the defined methods, etc. and (ii) process all external references defined in the 
bytecode.


\section{Client side}\label{sect:clientside}

\subsection{Eclipse Integrated Development Environment~\cite{Eclipseproject}}
The Eclipse Project is an open source software development
project dedicated to providing a robust, full-featured, commercial-quality,
industry platform for the development of highly integrated tools. It was
developed by IBM from 1999, and a few months after the first version was
shipped, IBM donated the source code to the Eclipse Foundation.

The Eclipse project consists of many subprojects, the most important being the
Eclipse Platform, that defines the set of frameworks and common services that
collectively make up ,,integ\-ration-ware'' required to support a comprehensive
tool integration platform. These services and frameworks represent the common
facilities required by most tool builders, including a standard workbench user
interface and project model for managing resources, portable native widget and
user interface libraries, automatic resource delta management for incremental
compilers and builders, language-independent debug infrastructure, and
infrastructure for distributed multi-user versioned resource management.

The Eclipse Platform has an easily-extendable modular architecture, where all
functionality is achieved by plugins, running over a low-level core called
Platform Runtime. This runtime core is only responsible for loading and
connecting the available plugins, every other functionality, such as the
editors, views, project management, is handled by plugins. The plugins bundled
with Eclipse Platform include general user interface components, a common help
system for all Eclipse components, project management and team work support.

\subsection{Eclipse Java Development Tools~\cite{JDT}}
The Eclipse Java Development Tools (JDT) project contributes a set of
plugins that add the capabilities of a full-featured Java IDE to the Eclipse
platform. The JDT plugins provide APIs so that they can themselves be further
extended by other tool builders. The JDT plugins are categorized into five main
projects: Annotation Processing (APT), Core, Debug, Text, and User Interface
(UI).

The most important part -- in the context of this paper -- is the Core project
which implements all the necessary non-UI infrastructure for Java application
development on the Eclipse platform. It contains an incremental Java builder,
support for code assist, a searching facility and a \emph{Java Model}, which
provides API for navigating the Java element tree. This lets the contributors
access the structure and the changes of the Java applications loaded into the
workspace.

\subsection{Eclipse Modeling Framework~\cite{EMF}}
The Eclipse Modeling Framework (EMF) is a Java framework and code
generation facility for building tools and other applications based on a
structured model. EMF is also maintained by the Eclipse Foundation within the
scope of Eclipse Tools Project. EMF started out as an implementation of the OMG
Meta Object Facility (MOF) specification, and evolved into a highly efficient
product for model-based software design.

EMF provides automated code generation and tooling (e.g.~notification,
persistence, editor) for Java representation of models. EMF models consist of an
(acyclic) containment hierarchy of model elements (\emph{EObjects}) with
crossreferences -- some of which may only be traversed by programs in one
direction (unidirectional references). Additionally, each object has a number of
attributes (primitive data values). Models are stored in \emph{EResources} (e.g.
files), and interrelated resources are grouped into \emph{EResourceSets}.

EMF provides the foundation for interoperability with other EMF-based tools and
applications. EMF has a built-in serialization facility, which enables the developer to save
(and load) in-stances of the model into industry-standard XMI format. EMF also
provides a notational and persistence mechanism, on top of which
model-manipulating tools can easily be constructed.

\subsubsection{Ecore Metamodeling~\cite{MODELS10}}
EMF uses \emph{Ecore} metamodels to describe the abstract syntax of a modeling
language. The main elements of Ecore are the following: \emph{EClass}
(represented graphically by a rectangle on \autoref{fig:cp3model.pdf}),
\emph{EAttribute} (entries in the rectangle) and \emph{EReference} (depicted as
edges). EClasses define the types of EObjects, enumerating EAttributes to
specify attribute types of class instances and EReferences to define association
types to other EObjects. Some EReferences additionally imply containment
(graphically represented by a diamond).
Unidirectional references are represented by arrows. Both ends of an association
may have a multiplicity constraint attached to them, which declares the number
of objects that, at run-time, may participate in an association. The most
typical multiplicity constraints are i) the at-most-one (0..1), and (ii) the
arbitrary (denoted by *). Inheritance may be defined between classes (depicted
by a hollow arrow), which means that the inherited class has all the properties
its parent has, and its instances are also instances of the ancestor class, but
it may further define some extra features.


\subsubsection{EMF-IncQuery~\cite{icmt2011}} 

% Problem to solve.
When working with models, it is quite common task to query the model for
validating certain properties or searching for interesting parts. EMF-IncQuery
\cite{EMFIncQuery} provides a solution for this problem: It is a framework to
execute fast model queries over EMF models. It is actively developed at Budapest
University of Technology and Economics. The current stable version  (0.6.8) 
proved its usability through industrial use-cases
and university researches.

% Core.
The core of the framework is a query evaluator engine built on top of graph
pattern matching engine using RETE \cite{RETE} algorithm adapted from expert
systems to facilitate the efficient storage and retrieval of partial views of
graph-like models. In a nutshell, RETE maintains a hierarchical query data
structure on top of the model which stores the result of sub-queries. On model
change, the event propagates through this data structure leaving the unmodified
part of the model untouched. This results in fast, near zero response time and
size-independence on small model changes. In return, the model and the query
structure has to be loaded into the memory, which can be a significant resource
expense.

% API.
To access the capabilities of the core, an easy-to-use, type safe API is
defined. Using the API, EMF resources and object hierarchies can be loaded and
queried incrementally. In addition certain extensions -- such as the validation
framework -- can be attached.

% Query lanuage.
Along the API, a complete query language is defined. It provides a declarative
way to express the queries over the EMF model in the form of graph patterns.
With the language the user can express combined queries, negative patterns,
checking property conditions, simple calculations, calculate disjunctions and
transitive closures, etc.~on top of the models.

% Tooling.
The framework contains UI tooling which helps the users to effectively develop
test and integrate queries into their solution. The first element of the tooling
is the rich XText~\cite{XtextWeb} based editor for the query language which aids writing well
formed queries providing content assist, error markings and such. The next part
of the tooling is the ability to load EMF models and execute the queries on them
as the user writes them giving visual feedback about the result.
The last important part is the code generation. The tooling dynamically
generates the source code which contains the programmatic equivalent of the
model queries. The users can integrate this code out of the box in Eclipse
plugins as well as headless applications to execute queries and get back the
results from the source code level.



